### 2022-11-26 manuscript.Rmd
minor formatting fix
Small rephrasing

### 1.2.1	Statistical Power to Detect a Specific Effect of Interest
Central to many studies that employ SEM is the question of whether a specific parameter in a model (e.g., strength of a specific correlation or the slope parameter in a latent regression model) is different from a given value [e.g. @canada_2021]. Imagine that a researcher aims to test whether the correlation between two latent constructs significantly differs from zero. To this end, they would compare the model fit between one model in which the correlation parameter is freely estimated and another model in which it is fixed to zero [@satorra_1985]. The corresponding null hypothesis in this case states that the parameter restrictions hold in the population. The difference in model fit will follow a $\chi^2$-distribution with degrees-of-freedom (*df*) equal to the difference of freely estimated parameters between the two models, if the null hypothesis is true [@neale_2000, in this example *df* = 1, because only the correlation between the two latent factors is fixed in the restricted model]. If restricting the parameter of interest results in a significantly poorer model fit, this would suggest that the respective parameter indeed significantly differs from zero.
Following the logic outlined above, researchers can determine the statistical power for detecting a target effect in an SEM framework before conducting the study. A key aspect for this a-priori power calculation hinges on translating the differences in a specific parameter estimate into an effect size. Such translation can be achieved by investigating the discrepancy between the model-implied variance-covariance matrices associated with (1) the population model including the “true” parameter values and (2) the hypothesized model. The discrepancy between both matrices is quantified based on a specific fit function [for details on fit functions see @bollen_1989]. Recently, several user friendly tools have emerged that can guide modelers in quantifying effect sizes of a target effect to estimate statistical power in SEM [e.g. the R packages semPower @jobst_2021; or power4SEM @jak_2020 for analytical approaches; and the Shiny app pwrSEM @wang_2021; the interactive study planner tool LIFESPAN @brandmaier_2015; or the R package simsem  @pornprasertmanit_2021 for simulation-based approaches]. As the investigation of a target effect practically translates into the comparison of nested models, researchers can in these cases analytically determine the statistical power and use this information for decisions on required sample size [@satorra_1985].

@@

First, we showed that detecting differences in model fit requires increasingly many participants the higher the inter-correlation of the factors turns out practically. Therefore, it is warranted to consider the actual separability of the constructs in question to avoid underpowered sampling plans. We acknowledge that committing to specific parameter values, such as factor inter-correlations, can be difficult when setting up the population models for the simulations. Thus, we recommend researchers to consult existing literature in order to make informed assumptions on the relationship among different latent factors. For cases in which no or little evidence exists on the relationship between the constructs under investigation, we recommend to enter these inter-correlations into the simulation as parameter of interest (as demonstrated in the example of this paper) and to then carefully evaluate what level of resolution is required when separating inter-related factors depending on practical relevance.

Second, the results on the loadings of manifest variables stresses the importance of employing indicators that capture the latent construct with high reliability [@zuo_2019]. Considering the issue of reliability is a crucial step in experimental design, as the reliability of the indicators heavily influences the practical ability to identify the targeted latent construct with sufficient precision and differentiate between competing models. Again, committing to population values can be difficult, especially if there is no prior work from which to draw values. In this case, we still advice to estimate plausible values from related scientific areas. For instance, selecting factor loading strength in the tri-partite model is not straightforward. However, we can do a literature search on similar models and take the reported test- reliability (e.g., a test-retest) or construct reliability (e.g., coefficient alpha) and translate this back into factor loadings strength and residual loadings [see @brandmaier_2018 for examples on how to parametrize longitudinal models from minimal available information]. If information about test reliability is available, we set the loadings to the square-root of the reliability to take this prior information into account. Then, this yields the following construct reliability for our design (assuming identical loadings across *k* items):

>>>>

### 1.2.1	Statistical Power to Detect a Specific Effect of Interest
Central to many studies that employ SEM is the question of whether a specific parameter in a model (e.g., strength of a specific correlation or the slope parameter in a latent regression model) is different from a given value [e.g. @canada_2021]. Imagine that a researcher aims to test whether the correlation between two latent constructs significantly differs from zero. To this end, they would compare the model fit between one model in which the correlation parameter is freely estimated and another model in which it is fixed to zero [@satorra_1985]. The corresponding null hypothesis in this case states that the parameter restrictions hold in the population. The difference in model fit will follow a $\chi^2$-distribution with degrees-of-freedom (*df*) equal to the difference of freely estimated parameters between the two models, if the null hypothesis is true [@neale_2000, in this example *df* = 1, because only the correlation between the two latent factors is fixed in the restricted model]. If restricting the parameter of interest results in a significantly poorer model fit, this would suggest that the respective parameter indeed significantly differs from zero.  
Following the logic outlined above, researchers can determine the statistical power for detecting a target effect in an SEM framework before conducting the study. A key aspect for this a-priori power calculation hinges on translating the differences in a specific parameter estimate into an effect size. Such translation can be achieved by investigating the discrepancy between the model-implied variance-covariance matrices associated with (1) the population model including the “true” parameter values and (2) the hypothesized model. The discrepancy between both matrices is quantified based on a specific fit function [for details on fit functions see @bollen_1989]. As the investigation of a target effect practically translates into the comparison of nested models, researchers can in these cases analytically determine the statistical power and use this information for decisions on required sample size [@satorra_1985]. Recently, several user friendly tools have emerged that can guide modelers in deriving estimates for statistical power to estimate statistical power in SEM in such cases [e.g. the R packages semPower @jobst_2021; or power4SEM @jak_2020 for analytical approaches; and the Shiny app pwrSEM @wang_2021; the interactive study planner tool LIFESPAN @brandmaier_2015; or the R package simsem  @pornprasertmanit_2021 for simulation-based approaches].

@@

First, we showed that detecting differences in model fit requires increasingly many participants the higher the inter-correlation of the factors turns out practically. Therefore, it is warranted to consider the actual separability of the constructs in question to avoid underpowered sampling plans. We acknowledge that committing to specific parameter values, such as factor inter-correlations, can be difficult when setting up the population models for the simulations. Thus, we recommend researchers to consult existing literature in order to make informed assumptions on the relationship among different latent factors. For cases in which no or little evidence exists on the relationship between the constructs under investigation, we recommend to enter these inter-correlations into the simulation as parameter of interest (as demonstrated in the example of this paper) and to then carefully evaluate what level of resolution is required when separating inter-related factors depending on practical relevance.  
Second, the results on the loadings of manifest variables stresses the importance of employing indicators that capture the latent construct with high reliability [@zuo_2019]. Considering the issue of reliability is a crucial step in experimental design, as the reliability of the indicators heavily influences the practical ability to identify the targeted latent construct with sufficient precision and differentiate between competing models. Again, committing to population values can be difficult, especially if there is no prior work from which to draw values. In this case, we still advice to estimate plausible values from related scientific areas. For instance, selecting factor loading strength in the tri-partite model is not straightforward. However, we can do a literature search on similar models and take the reported test- reliability (e.g., a test-retest) or construct reliability (e.g., coefficient alpha) and translate this back into factor loadings strength and residual loadings [see @brandmaier_2018 for examples on how to parametrize longitudinal models from minimal available information]. If information about test reliability is available, we set the loadings to the square-root of the reliability to take this prior information into account. Then, this yields the following construct reliability for our design (assuming identical loadings across *k* items):
